Tutkielman otsikko

### 2020-02-03 gradu/msc_english_contents.tex
Adds more text

\chapter{Basic Concepts}
Variable-byte encoding \citep{Wil99} is a method of compressing integers. It splits an integer into blocks of $b$ bits and then encodes
the block as $b + c$. $c$ is a bit denoting whether $b$ has the most significant bits of the number or not. (example?) A good dataset for variable-byte encoding
is a set with both very small and very large numbers, since the compression comes from omitting the leading zeros in the bit array. This method 
loses in compression performance to other methods \citep{Bri09}, but decoding is fast.

Rank and select are two functions that work on bit arrays. Rank(i) gives the sum of 1 bits from the beginning of the bit array and select(i) gives
the index of ith 1 bit in the bit array. Both functions work in constant time (citation?) and they require only a few percents of extra space over
the data. The extra bits $c$ added by variable-byte encoding conveniently create a bit array, where 1's represent the endings of numbers. An effective
version of random access has already been introduced \citep{Bri09}. It uses one rank call for each block after the first which makes the fetching very
fast if the number in question fits to one block.

An effective version with select is also possible. The index of the block where ith number starts is obtainable with select(i-1)+1, when i > 0 and 0 when i=0.
By separating the $c$ bit array and $b$ block array, $b$ now contains variable-byte integers in readable form. Another upside is that functions 
next(i) and previous(i) are conveniently available without any extra calculation and the runtime is not dependent on the size of the value.

@@

\chapter{Algorithm}
 - modifications to basic implementation

\chapter{Results}
 - comparison to basic implementation + Bri09
 - 
\chapter{Conclusion}
 - here

@@

\begin{table}
\centering
\caption{Results with 100k entries (in milliseconds).\label{table:results}}
\begin{tabular}{l||l c c c c r} 
Experiment & 128 & 256 & 32768 & 65536 & $2^{24}$ & $2^{32} -1$\\ 
\hline \hline 
$7bit VByte encoding$ & 34.97 & 49 & 53.04 & 52.18 & 53.08 & 76.21\\
$8bit VByte encoding$ & 33.57 & 32.47 & 42.96 & 43.11 & 46.15 & 65.14\\
$7bit VByte encoding with array$ & 33.39 & 46.85 & 51.24 & 49.03 & 48.93 & 66.84 \\
$8bit VByte encoding with array$ & 32.52 & 31.88 & 41.54 & 39.94 & 41.15 & 52.86 \\

>>>>

\chapter{Variable-byte encoding}

Variable-byte encoding \citep{Wil99} (VB) is a method of compressing integers via omitting leading zero bits. A good data set for 
VB encoding is a list of different sized numbers. As an example, an inverted index used by document search engine stores word frequencies and 
locations in documents. VB saves space here because index numbers need to support large values, but are usually significantly smaller. 
It is also proven to improve search speed, because less bytes need to be read from a hard drive \citep{Sch02}.

VB splits an integer into blocks of $b$ bits. A continuation bit is added to the block to form a chunk $c$ and then chunk non-empty chunks are stored.
The continuation bit is set to 1 for the last block of the integer. During decoding, chunks are read and their blocks are joined until a continuation 
bit is set to 1. An example implementation is showcased in Figure~\ref{fig:vbyte}.

\begin{figure}[h!] 
\centering 
blblblblb
testataan
tätä
\caption{VB encoding and decoding.\label{fig:vbyte}}
\end{figure}

Small lengths of $c$ can yield better
compression rate at the cost of more bit manipulation, while longer chunks need less bit manipulation and offer less effective compression. 
Generally chunk length of 8 has been used because it gives a good average and handling bytes is convenient \citep{IRbook}.

bl 
It splits an integer into blocks of $b$ bits and then encodes
the block as $b + c$ where $c$ is a bit denoting whether $b$ has the most significant bits of the number or not.

This method loses in compression performance to other methods \citep{Bri09}, but decoding is fast.

\chapter{Directly addressable codes?}
Rank and select are two functions that work on bit arrays. Rank(i) gives the sum of 1 bits from the beginning of the bit array and select(i) gives
the index of ith 1 bit in the bit array. Both functions work in constant time (citation?) and they require only a few percents of extra space over
the data. The extra bits $c$ added by variable-byte encoding conveniently create a bit array, where 1's represent the endings of numbers. An effective
version of random access has already been introduced \citep{Bri09}.

Random access with select query is also possible.  By separating the $c$ bit array and $b$ block array, $b$ contains variable-byte integers 
in readable form. Another upside is that functions next(i) and previous(i) are conveniently available. Rank implementation has

@@

\chapter{Algorithm}

 - modifications to basic implementation

\chapter{Results}
 - comparison to basic implementation + Bri09
\begin{table}
\centering
\caption{Results with 100k entries (in milliseconds).\label{table:results}}
\begin{tabular}{l||l c c c c r} 
Experiment & 128 & 256 & 32768 & 65536 & $2^{24}$ & $2^{32} -1$\\ 
\hline \hline 
$7bit VByte encoding$ & 34.97 & 49 & 53.04 & 52.18 & 53.08 & 76.21\\
$8bit VByte encoding$ & 33.57 & 32.47 & 42.96 & 43.11 & 46.15 & 65.14\\
$7bit VByte encoding with array$ & 33.39 & 46.85 & 51.24 & 49.03 & 48.93 & 66.84 \\
$8bit VByte encoding with array$ & 32.52 & 31.88 & 41.54 & 39.94 & 41.15 & 52.86 \\

\chapter{Conclusion}
 - here

@@
